Designing and Developing a Personalised Recommender System

### 2021-02-11 paper/paper.tex
changed demo user IDs to better ones

\subsection{Data preparation and feature selection}
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
The list of reviews pertaining to these restaurants was then filtered down
I also split by nearest city, and filtered only useful reviews
This was done using the $categories$ attribute of the $business$ JSON file.

The business data was filtered down to businesses that contain 'Restaurants' as one of their listed 
categories
getIDs
filterReviews
lastCheckin
recentCheckin
toCSV
CovidDuplicates
getUserIDs
mostActiveUsers
numberRestaurantReviews
mostReviews
filterMostActiveReviews
uniqueRestaurants
dropText
splitCities
splitIDs
filterUsefulReviews
filterRecentReviews
splitReviews
getCategories
formatCovidFeatures
closedCovid

The ratings data was 
(20,565 restaurants)
Prepare data
Of the 209,393 businesses in the dataset, 168,903 are still open. 
There are 43,965 restaurants which are open and contain 'Restaurants' as one of the categories. 
I then picked the top 12 most popular categories of restaurant: American, Mexican, Italian, Chinese, 
Seafood, Japanese, Canadian, Mediterranean, Indian, Thai, Middle Eastern, Vietnamese (in that order). 
When filtered for restaurants which fall into at least one of these categories, 
20,565 had at least one check-in in 2019, but 333 of these were temporarily closed due to covid but should be 
open now, and 1 is still closed until June 2021. 
19,862 restaurants had a review written about them in 2019, there were 503,979 reviews written in 2019 about these restaurants. 
There are 2,956,316 reviews of these restaurants. 
481,073 tips for these restaurants. 
1,968,703 users. 
None of these restaurants are temporarily closed according to the covid features data. 
20,573 restaurants listed with covid features, 4 restaurants which make up 8 duplicate entries 
(4 have 3 entries), the only difference between their entries being in the 
"Virtual Services Offered" section. 
Allow to pick user ID of 10 users with most reviews made, to avoid cold start problem
4 users have over 10,000 reviews, the most is 14,455 (Victor, 8k3aO-mPeyhbR5HUucA5aA). 
I actually want to select the most reviews of the places I am accepting, restaurants. 
When taking the most reviews of restaurants in teh dataset, a threshold of 600 reviews gives a list of 10 users, 
with the most being 1747. 
I then checked each of these for how many different restaurants they had reviewed and the greatest number of times 
they had reviewed a single restaurant to check that none of them were just users who reviewed a small number of 
restaurants a huge amount of times each. 
All 10 users passed this criteria, with even the least active user having reviewed 496 unique restaurants, which will 
certainly avoid any cold-start problems. 
I split the businesses into 10 groups, by the city which they are closest to, I did this by using latitude and 
longitude rather than State or City since this does not give actual distance and restaurants close to a border 
could then be mis-categorised. 
Since there are 4,223,201 reviews of restaurants I filtered them by removing any which had zero 'useful' votes. 
When filtered to reviews with at least 1 'useful' vote, there are 1,735,732 and when at least 2 votes, 836,929
At least 3 votes, 462,412, at least 5 votes 190,902. 
51,737 with at least 10 votes

>>>>

\subsection{Data preparation and feature selection}
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I chose to use the top 12 categories because it's enough to provide a wide variety of options to the user without 
being so many that the user is overwhelmed by the amount. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
All other datasets were filtered to only data pertaining to these restaurants, and the reviews were further filtered 
to only those which had 5 or more 'useful' votes by other users. 
This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) and structured as follows: 
($User\_ID$, $Business\_ID$, $Rating$).

20,565 had at least one check-in in 2019, but 333 of these were temporarily closed due to covid but should be 
open now, and 1 is still closed until June 2021. 
19,862 restaurants had a review written about them in 2019, there were 503,979 reviews written in 2019 about these restaurants. 
There are 2,956,316 reviews of these restaurants. 
481,073 tips for these restaurants. 
1,968,703 users. 
None of these restaurants are temporarily closed according to the covid features data. 
20,573 restaurants listed with covid features, 4 restaurants which make up 8 duplicate entries 
(4 have 3 entries), the only difference between their entries being in the 
"Virtual Services Offered" section. 
Allow to pick user ID of 10 users with most reviews made, to avoid cold start problem
4 users have over 10,000 reviews, the most is 14,455 (Victor, 8k3aO-mPeyhbR5HUucA5aA). 
I actually want to select the most reviews of the places I am accepting, restaurants. 
When taking the most reviews of restaurants in teh dataset, a threshold of 600 reviews gives a list of 10 users, 
with the most being 1747. 
I then checked each of these for how many different restaurants they had reviewed and the greatest number of times 
they had reviewed a single restaurant to check that none of them were just users who reviewed a small number of 
restaurants a huge amount of times each. 
All 10 users passed this criteria, with even the least active user having reviewed 496 unique restaurants, which will 
certainly avoid any cold-start problems. 
I split the businesses into 10 groups, by the city which they are closest to, I did this by using latitude and 
longitude rather than State or City since this does not give actual distance and restaurants close to a border 
could then be mis-categorised. 
Since there are 4,223,201 reviews of restaurants I filtered them by removing any which had zero 'useful' votes. 
When filtered to reviews with at least 1 'useful' vote, there are 1,735,732 and when at least 2 votes, 836,929
At least 3 votes, 462,412, at least 5 votes 190,902. 
51,737 with at least 10 votes
